Finite-size scaling of free energy in the dimer model on a hexagonal domain

### 2019-08-03 AffineLieAlgebras/MQFT2018/paper.tex
New version in the arXiv

\usepackage{amsmath,amssymb,amsthm,amsfonts}

@@

We consider a particular case of the dimer model on a hexagonal domain of hexagonal lattice, that
can be seen as a ``pile of cubes in the corner of the box''. The energy of configuration is the
total number of cubes. For this particular case we use an exact combinatorial formula for the
partition function to derive the expressions for scaling limit of free energy and first three terms
of the finite-size corrections. We show that the first term is identically zero. The second term
which encapsulates logarithmic dependence on the mesh size is connected with the central charge of
the effective field theory and with the geometry of limit shape. The third term is
geometry-dependent and is written explicitly in terms of elementary functions.

We then consider the box of infinite height and the case of coordinate-dependent Boltzmann weights.
In these cases we are also able to write explicit expressions for finite-size corrections. We
demonstrate the universality of the logarithmic term.

@@

\section{Model definition}
\label{sec:model-definition}
The configurations of the dimer model are perfect matchings (sets of non-touching edges, covering
all the vertices) on some graph ${\cal G}$ with some choice of weights $\omega(e)$ on the edges. The
model is solvable on the bipartite graphs, i.e. the partition function can be computed if the
weights are introduced in such a way that for each face bounded by 0 mod 4 edges there is an odd
number of negative edge weights and each face with 2 mod 4 edges has an even number of of negative
edge weights. Then the signs of the edge weights form a so-called ``Kasteleyn orientation''on graph,
the weighting is called ``Kasteleyn weighting'' \cite{kenyon2001dominos,kenyon2009lectures}.

@@

The weight of a particular configuration is given by the exponent of the volume of all cubes or by a
sum of the height function values:
\begin{equation*}
  \label{eq:10}
  E[conf]=\sum_{i,j} h_{ij}=\mathrm{Volume}
\end{equation*}
We set Boltzman constant equal to 1 and choose system of units in such a way that there is no
coupling constant in the expression for energy. Then the partition function is
\begin{equation*}
  \label{eq:14}
  Z=\sum_{conf} e^{-\frac{E[conf]}{T}}=\sum_{conf}q^{\mathrm{Volume}[conf]}, 
\end{equation*}
where $q=\exp\left(-1/T\right)$ .

@@

\section{Physical meaning of the expansion coefficients}
\label{sec:accur-expans-phys}

@@

In the scaling limit $\varepsilon\to 0$ so called ``limit shape phenomenon''~\cite{1998math......1068J,cohn1998shape}
appears in the dimer model. The areas around the corners of the domain are ``frozen'' with height
function values being fixed. An analytical ``Arctic curve'' delimits frozen regions from the region
where the behavior is described by the effective free field theory
\cite{kenyon2009lectures,kenyon2008height,kenyon2006dimers}.

The behavior (\ref{eq:26}) of the logarithm of the partition function is generic in two-dimensional
models \cite{cardy1988finite}.

First two terms $f_{0}$ and $f_{1}$ are interpreted as a bulk and
boundary free energies in the corresponding field theory. Since we have $f_{1}=0$, we can conclude
that boundary tension is zero.
The value of the first term $f_{0}$ can be rewritten using polylogarithm functions as

The term proportional to the logarithm of the scale $\varepsilon$ is also universal
\cite{cardy1988finite} and should appear in all two-dimensional theories with boundary. In the paper
\cite{cardy1988finite} it was argued that on a manifold of a characteristic length $L$ with a smooth
boundary such a term must have the following form:
\begin{equation}
  \label{eq:31}
  \delta F = -\frac{1}{6}{\bf c} \chi \ln L,
\end{equation}
where ${\bf c}$ is central charge of the effective field theory and $\chi$ is the Euler characteristic of
the manifold
\begin{equation}
  \label{eq:32}
  \chi=2-2 h-b,
\end{equation}
where $h$ is the number of handles and $b$ is the number of boundaries.

Since the non-frozen domain is delimited by a smooth boundary, we have $\chi=1$ and we interpret
$2(ab+bc+ca)$ as a volume of the domain, thus
\begin{equation}
  \label{eq:33}
  \delta F = \left(2(ab+bc+ca)\right) f_{2}\ln\varepsilon,
\end{equation}
which together with formula \eqref{eq:30} suggests the value of central charge
${\bf c}=\frac{1}{2}$. This value is in agreement with the identification of the dimer model with
free fermions in the paper \cite{dijkgraaf2009dimer}.

The last term $f_{3}$ depends only on the shape of the domain through $a,b,c$ with a universal
contribution that is equal to the integral of a function $e^{-z}\frac{\xi(z)-\xi(0)}{z}$. In the
next sections we show that this constant appear even for a different choice of Kasteleyn weighting,
due to the treatment of logarithmic divergency. In the last section work we consider coordinate
dependent values of $q$ and more complex geometries can be treated similarly as was done in the
paper \cite{okounkov2007random}.

@@

Now we need to consider other terms in the expression \eqref{eq:55}. For $f^{+}$ we have
\begin{equation}
  \label{eq:67}
  f^{+}=\frac{1}{ab}\sum_{n=1}^{\infty}\frac{1}{n} e^{-n\varepsilon\varphi(b-a)} \varepsilon
  \frac{1-e^{-nbr_{-}'(0)}}{1-e^{-n\varepsilon r_{-}'(0)}}\varepsilon\sum_{j=0}^{M-1}\Psi^{+}_{n}(j\varepsilon).
\end{equation}
The sum over $j$ can be approximated by an integral using Euler-Maclaurin formula as
\begin{equation}
  \label{eq:68}
  \varepsilon\sum_{j=0}^{M-1}\Psi^{+}_{n}(j\varepsilon) = \int_{0}^{a}\Psi^{+}_{n}(x) \dx -
  \frac{\varepsilon}{2}\left(\Psi^{+}_{n}(a)-\Psi^{+}_{n}(0)\right)+
  \frac{\varepsilon^{2}}{12}\left(\Psi^{+'}_{n}(a)-\Psi^{+}_{n}(0)\right)+
  \mathcal{O}\left(\varepsilon^{m}\Psi^{+(m-1)}_{n}\right).
\end{equation}
Since $r(x)$ is an analytic function and $r(0)=0$ we can write $r(x)=xr'(0)+x^{2}\zeta(x)$, so 
\begin{equation}
  \label{eq:83}
\Psi_{n}(x)=e^{-nr'(0)x}\left(e^{-n\left(r(x)-xr'(0)\right)}-1\right)=e^{-nxr'(0)}\left(e^{-nx^{2}\zeta(x)}-1\right)  
\end{equation}
and we see that $\Psi_{n}(0)=0$ and $\Psi_{n}'(0)=0$.
Than in the limit $n\to\infty$ we have
\begin{equation}
  \label{eq:69}
  \int_{0}^{c}\Psi_{n}(x)
  \dx \xrightarrow[n\to\infty]{}\int_{0}^{c}e^{-nxr'(0)}\left(-nx^{2}\zeta(x)+\dots\right)\dx\sim \frac{1}{n^{2}}.
\end{equation}
The derivatives of $\Psi$ behave as $\Psi_{n}^{(m)}(0)\sim n^{m-1}$, since the derivatives should
fall on the bracket in the expression \eqref{eq:83} for the value to be non-zero, and then extra
power of $x$ is produced, which in turn should be differentiated. So the contribution from
$\varepsilon^{m}\Psi_{n}^{(m-1)}(0)$ to $f^{+}$ gives us
\begin{equation}
  \label{eq:70}
  \sum_{n=1}^{\infty} \frac{1}{n} e^{-n\varepsilon\varphi(b-a)}
  \frac{1-e^{-nbr_{-}'(0)}}{1-e^{-n\varepsilon r_{-}'(0)}} \varepsilon^{m+1}\Psi_{n}^{(m-1)}(0)\sim
  \sum_{n=1}^{\infty} \frac{1}{n^{4}} e^{-n\varepsilon\varphi(b-a)}
  \frac{1-e^{-nbr_{-}'(0)}}{1-e^{-n\varepsilon r_{-}'(0)}} (\varepsilon n )^{m+1}.
\end{equation}
For the second derivative in $\varepsilon$ we thus see that main contribution behaves as
$\sum_{n=1}\frac{1}{n^{2}}e^{-n\varepsilon\varphi(b-a)}
  (n\varepsilon)^{m+1}\xrightarrow[\varepsilon\to 0]{}0$
if $m\geq 3$. For the terms with $m=0,1,2$ we need to
take the limit $\varepsilon\to 0$. Doing so, we obtain
\begin{equation}
  \label{eq:71}
  f^{+}_{3}\equiv\left.\left(\frac{d}{d\varepsilon}\right)^{2} f^{+}\right|_{\varepsilon=0} =
  \left.\left(\frac{d}{d\varepsilon}\right)^{2}\frac{1}{ab}\sum_{n=1}^{\infty} \frac{1}{n}
  e^{-n\varepsilon\varphi(b-a)} \frac{1-e^{-nbr_{-}'(0)}}{1-e^{-n\varepsilon r_{-}'(0)}}
  \varepsilon\left(\int_{0}^{a} \Psi_{n}^{+}(x) \dx -\varepsilon \Psi_{n}^{+}(a) +\varepsilon^{2}\Psi^{+'}_{n}(0)\right)\right|_{\varepsilon=0}.
\end{equation}
Note that $\frac{\varepsilon}{1-\exp(-n\varepsilon r'_{-}(0))}$ is analytic in $\varepsilon$. 
The derivative can be computed explicitly, but the expression is too long to present it here.

The contribution of $f^{-}$ is computed in exactly the same way:
\begin{equation}
  \label{eq:72}
   f^{-}_{3}\equiv \left.\left(\frac{d}{d\varepsilon}\right)^{2} f^{-}\right|_{\varepsilon=0} =
  \left.\left(\frac{d}{d\varepsilon}\right)^{2}\frac{1}{ab}\sum_{n=1}^{\infty} \frac{1}{n}
  e^{-n\varepsilon\varphi(b-a)} \frac{1-e^{-nar_{+}'(0)}}{1-e^{-n\varepsilon r_{+}'(0)}}
 \varepsilon\left(\int_{0}^{b} \Psi_{n}^{-}(x) \dx -\varepsilon \Psi_{n}^{-}(a) +\varepsilon^{2}\Psi^{-'}_{n}(0)\right)\right|_{\varepsilon=0}.
\end{equation}
At last, we need to consider
$f^{\times}=\frac{1}{ab}\sum_{n=1}^{\infty}\frac{1}{n}e^{-n\varepsilon\varphi(b-a)}\varepsilon\sum_{i=0}^{N-1}\Psi_{n}^{-}(i\varepsilon)
\sum_{j=0}^{M-1}\Psi_{n}^{+}(j\varepsilon)$. Analyzing the contributions of the derivatives
$\Psi_{n}^{(m)}(0)$, similarly to the discussion before the formula \eqref{eq:70}, we see that
$\varepsilon^{m}\Psi_{n}^{(m-1)}(0)\sim \frac{1}{n^{2}}$. Thus
\[\sum_{n=1}^{\infty}\frac{1}{n}e^{-n\varepsilon\varphi(b-a)}\frac{1}{n^{2}}\varepsilon^{m}\Psi_{n}^{(m-1)}(0)\sim
\sum_{n=1}^{\infty}\frac{1}{n^{5}}e^{-n\varepsilon\varphi(b-a)}(\varepsilon n)^{m},\] and these terms
give zero contribution after the derivative $\left(\frac{d}{d\varepsilon}\right)^{2}$ for
$\varepsilon\to 0 $ and $m\geq 3$. The same holds for all the contributions with the derivatives of
$\Psi$ in the both sums. In the terms with $m<3$ we can just take a double derivative over
$\varepsilon$ and limit $\varepsilon\to 0$ and obtain
\begin{multline}
  \label{eq:73}
   f^{\times}_{3}\equiv \left.\left(\frac{d}{d\varepsilon}\right)^{2}
    f^{\times}\right|_{\varepsilon=0}=\left.\frac{1}{ab}\sum_{n=1}^{\infty}\frac{1}{n}
  \left(\frac{d}{d\varepsilon^{2}}\right)^{2}\left(\int_{0}^{b}\Psi^{-}_{n}(x)\dx
    -\frac{\varepsilon}{2}\Psi^{-}_{n}(b)+\frac{\varepsilon^{2}}{12}\Psi^{-'}_{n}(b)\right)\cdot\right.\\
\cdot\left.\left(\int_{0}^{a}\Psi^{+}_{n}(x)\dx-\frac{\varepsilon}{2}\Psi^{+}_{n}(a)+
    \frac{\varepsilon^{2}}{12}\Psi^{+}_{n}(a)\right)\right|_{\varepsilon=0}.
\end{multline}
Again, this expression is easily computed explicitly, but is too cumbersome to present it here.

@@

In the present paper we computed the asymptotic expansion of the free energy in the dimer model on a
hexagonal domain of the hexagonal lattice. We've discussed the physical meaning of the expansion
coefficients and argued that our results support the identification of the scaling behavior of the
dimer model with the free-fermion field theory.

In further work we will study the connection of the expansion coefficients with the spectral
properties of Dirac operator on the non-frozen domain and study the universality of our results. We
will also further discuss the description of the model by boundary conformal field theory.

\section*{Acknowledgments}
\label{sec:acknowledgements}
We are grateful to professor Nikolai Reshetikhin for his attention to this work. We thank Pavel Belov
for useful discussions and general support. We thank Mikhail Vyazovsky for pointing out a mistake in
the first version of the paper.

>>>>

\usepackage{amsmath,amssymb,amsthm,amsfonts}
%\synctex=1

@@

We consider a particular case of the dimer model on a hexagonal domain
of hexagonal lattice, that can be seen as a ``pile of cubes in the
corner of the box''. The energy of configuration is the total number
of cubes. For this particular case we use an exact combinatorial
formula for the partition function to derive the expressions for
scaling limit of free energy and first three terms of the finite-size
corrections. We show that the first term is identically zero. The
second term which encapsulates logarithmic dependence on the mesh size
is connected with the central charge of the effective field theory and
with the geometry of limit shape. The third term has a
geometry-dependent contribution that is written explicitly in terms of
elementary functions and a universal contribution, which is constant
in all the cases that we consider.

We then consider the box of infinite height and the case of
coordinate-dependent Boltzmann weights. In these cases we are also
able to write explicit expressions for finite-size corrections. We
demonstrate the universality of the logarithmic term and the constant
contribution to the third term.

@@

\section{Model definition}
\label{sec:model-definition}
The configurations of the dimer model are perfect matchings (sets of non-touching edges, covering
all the vertices) on some graph ${\cal G}$ with some choice of weights $\omega(e)$ on the edges. The
model is solvable on the bipartite graphs, i.e. the partition function can be computed if the
weights are introduced in such a way that for each face bounded by 0 mod 4 edges there is an odd
number of negative edge weights and each face with 2 mod 4 edges has an even number of of negative
edge weights. Then the signs of the edge weights form a so-called ``Kasteleyn orientation'' on graph,
the weighting is called ``Kasteleyn weighting'' \cite{kenyon2001dominos,kenyon2009lectures}.

@@

The weight of a particular configuration is given by the exponent of the volume of all cubes or by a
sum of the height function values:
\begin{equation*}
  \label{eq:10}
  E[conf]=\sum_{i,j} h_{ij}=\mathrm{Volume}
\end{equation*}
We set Boltzmann constant equal to 1 and choose system of units in such a way that there is no
coupling constant in the expression for energy. Then the partition function is
\begin{equation*}
  \label{eq:14}
  Z=\sum_{conf} e^{-\frac{E[conf]}{T}}=\sum_{conf}q^{\mathrm{Volume}[conf]}, 
\end{equation*}
where $q=\exp\left(-1/T\right)$ .

@@

\section{Limit shape and physical meaning of the expansion coefficients}
\label{sec:accur-expans-phys}

@@

The dimers in the scaling limit are in general described by a
free-fermion theory \cite{dijkgraaf2009dimer}, but on a hexagonal
lattice (and on the graphs that admit Kasteleyn orientation) the model
is reformulated in terms of height function, that can be seen as a
bosonization of the theory \cite{gogolin2004bosonization}.

In the scaling limit $\varepsilon\to 0$ so called ``limit shape
phenomenon''~\cite{1998math......1068J,cohn1998shape} appears in the
dimer model. The areas around the corners of the domain are ``frozen''
with height function values being fixed. An analytical ``Arctic
curve'' delimits frozen regions from the region where the behavior is
described by the effective free field theory
\cite{kenyon2009lectures,kenyon2008height,kenyon2006dimers}. Moreover,
the height function in the scaling limit tends to certain analytical
surface and fluctuations around this ``limit shape'' are described by
a gaussian free field theory in a curved background of ``limit
shape''. A confomal map can be used to send the internal region of the
``Arctic curve'' onto the upper half-plane.

The scaling behavior (\ref{eq:26}) of the logarithm of the partition function is generic in two-dimensional
models \cite{cardy1988finite}.

Thus we can interpret first two terms $f_{0}$ and $f_{1}$ as a bulk
and boundary free energies in the corresponding field theory. The term
$f_1$ corresponds to the contribution of the ``limit shape''. Since we
have a deterministic frozen boundary, we naturally have $f_{1}=0$.
%The value of the first term $f_{0}$ can be rewritten using polylogarithm functions as

The term proportional to the logarithm of the scale $\varepsilon$ is also universal
\cite{cardy1988finite} and should appear in all two-dimensional theories with boundary. In the paper
\cite{cardy1988finite} Cardy and Peschel argued that on a smooth manifold of a characteristic length $L$ with a smooth
boundary such a term must have the following form:
\begin{equation}
  \label{eq:31}
  \delta F = -\frac{1}{6}{\bf c} \chi \ln L,
\end{equation}
where ${\bf c}$ is central charge of the effective field theory and
$\chi$ is the Euler characteristic of the manifold $\chi=2-2 h-b$,
where $h$ is the number of handles and $b$ is the number of
boundaries. Naive use of the formula \eqref{eq:31} with our result
$\delta F=\frac{1}{12}$ would give ${\bf c}=\frac{1}{2}$, but it is
not correct, since we need to consider the field theory in the curved
background.

In the case of a corner singularity on the boundary with the angle
$\Theta$ the logarithmic contribution to the free energy is changed to
\begin{equation}
  \label{eq:86}
  \delta F = \frac{{\bf c}}{24}\left(\frac{\Theta}{\pi}-\frac{\pi}{\Theta}\right) \ln L.
\end{equation}
This formula was recently used by N. Allegra \cite{allegra2015exact} to show that the central charge
in the dimer model on the square lattice with a corner monomer is ${\bf c}=1$, contrary to some
previous claims of ${\bf c}=-2$.

In the case of manifold with conic singularity with the semi-angle
$\Theta$, the logarithmic contribution presented in \cite{cardy1988finite} is:
\begin{equation}
  \label{eq:91}
  \delta F = \frac{{\bf c}}{24}\left(\frac{\Theta}{\pi}-\frac{4\pi}{\Theta}\right) \ln L.
\end{equation}

In our case we interpret $2(ab+bc+ca)$ as a volume of the domain, thus
\begin{equation}
  \label{eq:33}
  \delta F = \left(2(ab+bc+ca)\right) f_{2}\ln\varepsilon.
\end{equation}

Using Cardy-Peschel formulas, presented above, $\delta F$ can be
computed from value of the central charge for Gaussian free field
${\bf c}=1$, the curvature of metric, induced by the limit shape and
conformal map from the internal region of the ``Arctic curve'' to the
upper half-plane. Such a computation can be used to check our result
on $f_2$ but is quite lengthy and will be presented in a separate
publication.

The last term $f_{3}$ depends only on the shape of the domain through
$a,b,c$ with a universal contribution that is equal to the integral of
a function $e^{-z}\frac{\xi(z)-\xi(0)}{z}$. In the next sections we
show that this constant appears due to the treatment of logarithmic
divergency even for the coordinate-dependent Boltzmann weights.

@@

Now we need to consider other terms in the expression \eqref{eq:55}. For $f^{+}$ we have
\begin{equation}
  \label{eq:67}
  f^{+}=\frac{1}{ab}\sum_{n=1}^{\infty}\frac{1}{n} e^{-n\varepsilon\varphi(b-a)} \varepsilon
  \frac{1-e^{-nbr_{-}'(0)}}{1-e^{-n\varepsilon r_{-}'(0)}}\varepsilon\sum_{j=0}^{M-1}\Psi^{+}_{n}(j\varepsilon).
\end{equation}
The sum over $j$ can be approximated by an integral using Euler-Maclaurin formula as
\begin{equation}
  \label{eq:68}
  \varepsilon\sum_{j=0}^{M-1}\Psi^{+}_{n}(j\varepsilon) = \int_{0}^{a}\Psi^{+}_{n}(x) \dx -
  \frac{\varepsilon}{2}\left(\Psi^{+}_{n}(a)-\Psi^{+}_{n}(0)\right)+
  \frac{\varepsilon^{2}}{12}\left(\Psi^{+'}_{n}(a)-\Psi^{+}_{n}(0)\right)+
  \mathcal{O}\left(\varepsilon^{m}\Psi^{+(m-1)}_{n}\right).
\end{equation}
Since $r(x)$ is an analytic function and $r(0)=0$ we can write $r(x)=xr'(0)+x^{2}\zeta(x)$, so 
\begin{equation}
  \label{eq:83}
\Psi_{n}(x)=e^{-nr'(0)x}\left(e^{-n\left(r(x)-xr'(0)\right)}-1\right)=e^{-nxr'(0)}\left(e^{-nx^{2}\zeta(x)}-1\right)  
\end{equation}
and we see that $\Psi_{n}(0)=0$ and $\Psi_{n}'(0)=0$.
Than in the limit $n\to\infty$ we have
\begin{equation}
  \label{eq:69}
  \int_{0}^{c}\Psi_{n}(x)
  \dx \xrightarrow[n\to\infty]{}\int_{0}^{c}e^{-nxr'(0)}\left(-nx^{2}\zeta(x)+\dots\right)\dx\sim \frac{1}{n^{2}}.
\end{equation}
The derivatives of $\Psi$ behave as $\Psi_{n}^{(m)}(0)\sim n^{m-1}$, since the derivatives should
fall on the bracket in the expression \eqref{eq:83} for the value to be non-zero, and then extra
power of $x$ is produced, which in turn should be differentiated. So the contribution from
$\varepsilon^{m}\Psi_{n}^{(m-1)}(0)$ to $f^{+}$ gives us
\begin{equation}
  \label{eq:70}
  \sum_{n=1}^{\infty} \frac{1}{n} e^{-n\varepsilon\varphi(b-a)}
  \frac{1-e^{-nbr_{-}'(0)}}{1-e^{-n\varepsilon r_{-}'(0)}} \varepsilon^{m+1}\Psi_{n}^{(m-1)}(0)\sim
  \sum_{n=1}^{\infty} \frac{1}{n^{4}} e^{-n\varepsilon\varphi(b-a)}
  \frac{1-e^{-nbr_{-}'(0)}}{1-e^{-n\varepsilon r_{-}'(0)}} (\varepsilon n )^{m+1}.
\end{equation}
For the second derivative in $\varepsilon$ we thus see that main contribution behaves as
$\sum_{n=1}\frac{1}{n^{2}}e^{-n\varepsilon\varphi(b-a)}
  (n\varepsilon)^{m+1}\xrightarrow[\varepsilon\to 0]{}0$
if $m\geq 3$. For the terms with $m=0,1,2$ we need to
take the limit $\varepsilon\to 0$. Doing so, we obtain
\begin{multline}
  \label{eq:71}
  f^{+}_{3}\equiv\left.\left(\frac{d}{d\varepsilon}\right)^{2} f^{+}\right|_{\varepsilon=0} =\\
  \left.\left(\frac{d}{d\varepsilon}\right)^{2}\frac{1}{ab}\sum_{n=1}^{\infty} \frac{1}{n}
  e^{-n\varepsilon\varphi(b-a)} \frac{1-e^{-nbr_{-}'(0)}}{1-e^{-n\varepsilon r_{-}'(0)}}
  \varepsilon\left(\int_{0}^{a} \Psi_{n}^{+}(x) \dx -\varepsilon \Psi_{n}^{+}(a) +\varepsilon^{2}\Psi^{+'}_{n}(0)\right)\right|_{\varepsilon=0}.
\end{multline}
Note that $\frac{\varepsilon}{1-\exp(-n\varepsilon r'_{-}(0))}$ is analytic in $\varepsilon$. 
The derivative can be computed explicitly, but the expression is too long to present it here.

The contribution of $f^{-}$ is computed in exactly the same way:
\begin{multline}
  \label{eq:72}
   f^{-}_{3}\equiv \left.\left(\frac{d}{d\varepsilon}\right)^{2} f^{-}\right|_{\varepsilon=0} =\\
  \left.\left(\frac{d}{d\varepsilon}\right)^{2}\frac{1}{ab}\sum_{n=1}^{\infty} \frac{1}{n}
  e^{-n\varepsilon\varphi(b-a)} \frac{1-e^{-nar_{+}'(0)}}{1-e^{-n\varepsilon r_{+}'(0)}}
 \varepsilon\left(\int_{0}^{b} \Psi_{n}^{-}(x) \dx -\varepsilon \Psi_{n}^{-}(a) +\varepsilon^{2}\Psi^{-'}_{n}(0)\right)\right|_{\varepsilon=0}.
\end{multline}
At last, we need to consider
$f^{\times}=\frac{1}{ab}\sum_{n=1}^{\infty}\frac{1}{n}e^{-n\varepsilon\varphi(b-a)}\varepsilon\sum_{i=0}^{N-1}\Psi_{n}^{-}(i\varepsilon)
\sum_{j=0}^{M-1}\Psi_{n}^{+}(j\varepsilon)$. Analyzing the contributions of the derivatives
$\Psi_{n}^{(m)}(0)$, similarly to the discussion before the formula \eqref{eq:70}, we see that
$\varepsilon^{m}\Psi_{n}^{(m-1)}(0)\sim \frac{1}{n^{2}}$. Thus
\[\sum_{n=1}^{\infty}\frac{1}{n}e^{-n\varepsilon\varphi(b-a)}\frac{1}{n^{2}}\varepsilon^{m}\Psi_{n}^{(m-1)}(0)\sim
\sum_{n=1}^{\infty}\frac{1}{n^{5}}e^{-n\varepsilon\varphi(b-a)}(\varepsilon n)^{m},\] and these terms
give zero contribution after the derivative $\left(\frac{d}{d\varepsilon}\right)^{2}$ for
$\varepsilon\to 0 $ and $m\geq 3$. The same holds for all the contributions with the derivatives of
$\Psi$ in the both sums. In the terms with $m<3$ we can just take a double derivative over
$\varepsilon$ and limit $\varepsilon\to 0$ and obtain
\begin{multline}
  \label{eq:73}
   f^{\times}_{3}\equiv \left.\left(\frac{d}{d\varepsilon}\right)^{2}
    f^{\times}\right|_{\varepsilon=0}=\left.\frac{1}{ab}\sum_{n=1}^{\infty}\frac{1}{n}
  \left(\frac{d}{d\varepsilon^{2}}\right)^{2}\left(\int_{0}^{b}\Psi^{-}_{n}(x)\dx
    -\frac{\varepsilon}{2}\Psi^{-}_{n}(b)+\frac{\varepsilon^{2}}{12}\Psi^{-'}_{n}(b)\right)\cdot\right.\\
\cdot\left.\left(\int_{0}^{a}\Psi^{+}_{n}(x)\dx-\frac{\varepsilon}{2}\Psi^{+}_{n}(a)+
    \frac{\varepsilon^{2}}{12}\Psi^{+}_{n}(a)\right)\right|_{\varepsilon=0}.
\end{multline}
Again, this expression is easily computed explicitly, but is too cumbersome to present it here.

@@

In the present paper we computed the asymptotic expansion of the free
energy in the dimer model on a hexagonal domain of the hexagonal
lattice, on a hexagon with a side of infinite length and in the model
with coordinate-dependent Boltzmann weights. We presented numerical
results supporting our computations. We've discussed the physical
meaning of the expansion coefficients and argued that our results
support the identification of the scaling behavior of the dimer model
with the Gaussian free field theory.

In further work we will study the connection of the expansion
coefficients with the spectral properties of Dirac operator on the
non-frozen domain and study the universality of our results. We will
also present the computations of the logarithmic correction to the free
energy from the geometry of the ``limit shape''.

\section*{Acknowledgments}
\label{sec:acknowledgements}
We are grateful to professor Nikolai Reshetikhin for his attention to
this work. We thank Pavel Belov for useful discussions and general
support. We thank Mikhail Vyazovsky for pointing out a mistake in the
first version of the paper. We thank Istvan Prause for the reference
\cite{allegra2015exact} and Vadim Gorin for a discussion on geometry
of the limit shape.
